---
title: "Homework 2"
author: "Dania Jafar"
date: "9/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

#Problem 1

```{r import_mr_tw}
trash_wheel_data = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  drop_na(., dumpster) %>% 
  mutate(sports_balls_new = as.integer(sports_balls))

data_2017 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(.) %>% 
  mutate(year = 2017)


data_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(.) %>% 
  mutate(year = 2018)

combined_data = union_all(data_2017, data_2018) %>% 
  mutate(month = month.name[month])

```

## Description of the dataframe 

* The number of observations in the 2017 Precipitation dataset are `r nrow(data_2017)` and the number of observations in the 2018 Precipitation dataset are `r nrow(data_2018)`

* Some of the key variables are month, total, and year

* The total precipitation in 2018 was `r sum(pull(data_2018, total))`

* The median number of sports balls in a dumpster in 2017 was `r median(pull(trash_wheel_data, sports_balls))`

```{r Problem 2}

polls = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = as.integer(month), 
  month = month.abb[month]) %>% 
janitor::clean_names() %>% 
mutate(month = tolower(month),
       year = as.numeric(year))
  
unemployment = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemployment_rate") %>% 
  select(year, month, unemployment_rate
)

snp = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(month = as.integer(month),
  month = tolower(month.name[month]),
  month = substr(month, 1, 3),
   year = as.numeric(year))

combined_data2 = 
  left_join(polls, snp, by = c("year", "month"))

combined_data3 =   
  left_join(combined_data2, unemployment, by = c("year", "month"))

combined_data3 = combined_data3 %>% 
 mutate (president = ifelse(prez_gop, "gop", "dem")) %>% 
  select(-day.x, -prez_dem, -prez_gop, -day.y)


## The pols data contained information related to the number of national politicians who are democratic or republican at any given time, the unemployment data contained information on unemployment rates for every month of a year,  the snp data set contains information related to Standard & Poor’s stock market index. The precipitation 2017 & precipitation 2018 data sets contain information on the amount of precipitation in those years. The Mr. Trash Wheel data set contains information on a vessel that removes trash from the Inner Harbor in Baltimore, Maryland.The also gives information on various items that were found in dumpsters from 2014-2016.


##Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).


```


```{r Problem 3}

baby_names = read_csv("./data/Popular_Baby_Names.csv") %>% 
janitor::clean_names() %>% 
distinct()




```




